---
title: "STA221"
author: "Neil Montgomery"
date: "Last edited: `r format(Sys.time(), '%Y-%m-%d %H:%M')`"
output: 
  beamer_presentation:
    keep_tex: TRUE
    incremental: TRUE
    df_print: tibble
    fig_caption: FALSE
classoption: aspectratio=169
header-includes:
- \newcommand{\ve}{\varepsilon}
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE,
                      dev='pdf', fig.width=5, fig.asp=0.618, fig.align='center')
options(tibble.width=70, scipen = 999, tibble.print_min=5, show.signif.stars = FALSE)
library(tidyverse)
library(readxl)
```


```{r}
bodyfat <- read_csv("Body_fat.csv")

options(scipen=6)
bf_wt <- bodyfat %>% 
  lm(`Pct BF` ~ Weight, data = .) 

bf_ht <- bodyfat %>% 
  lm(`Pct BF` ~ Height, data = .) 
```

## some test info

Bring a calculator. Suggested methodology:

1. Somehow get a calculator if you don't have one.

2. Bring it to the test.

\pause This is a two-step process. If you have one and don't bring it, you won't be able to use it. 

\pause I tease because there is always someone who forgets.

\pause Time and location: Monday, February 13, 15:20 to 16:50, in this room BA1160.

\pause There will be a formula sheet that comes with the test, along with $\chi^2$ and $t$ probability tables. I will publish these in advance so you will know what to expect.

# estimation and prediction with regression models

## estimate the mean response at a new $x$ value

Suppose you want to estimate the mean "response" at some new $x_\nu$ (may
or may not be one of the original $x$'s.)

\pause The *true value* for the mean response is:
$$\mu_\nu = \beta_0 + \beta_1 x_\nu$$

\pause What's the "obvious" best guess using the data?
$$\hat\mu_\nu = b_0 + b_1 x_\nu$$

## estimate the mean response---with confidence

A confidence interval will be as usual based on:
$$\frac{\hat\mu_\nu - \mu_\nu}{s.e.(\hat\mu_\nu - \mu_\nu)} \sim t_{n-2}$$

\pause The standard error of $\hat\mu_\nu - \mu_\nu$ is:
$$s_e\sqrt{\frac{1}{n} + \frac{(x_\nu - \overline x)^2}{S_{xx}}}$$

\pause So the 95\% C.I. for the mean response at $x_\nu$ will be:
$$\hat\mu_\nu \pm t^*_{n-2} s_e\sqrt{\frac{1}{n} + \frac{(x_\nu - \overline x)^2}{S_{xx}}}$$

## weight model example

Let's make a 95\% CI for the mean response at a weight of $x_\nu = 200$ pounds. Here's the R output:

```{r}
source("multiplot.R")
library(broom)
short_print_lm(summary(bf_wt))
```

## Weight at 200 example

```{r}
options(digits = 2)
```

\pause The estimate will be `r round(coefficients(bf_wt)[1], 3)` + `r round(coefficients(bf_wt)[2], 3)`(200) = `r predict(bf_wt, newdata = data_frame(Weight=200))`.

\pause $s_e = `r glance(bf_wt)$sigma`$

\pause We would need to be given $\overline x$, which in this case is the sample average of the `Weight` variable. This is: `r mean(bodyfat$Weight)`.

\pause We also need $S_{xx}$, which is `r (nrow(bodyfat)-1)*var(bodyfat$Weight)`. Could we have determined that from the output given?

\pause So the 95\% CI for the mean `Pct BF` at `Weight`=200 is:


$$`r predict(bf_wt, newdata = data_frame(Weight=200))` \pm `r qt(0.975, 248)` \cdot `r glance(bf_wt)$sigma`
\sqrt{\frac{1}{`r nrow(bodyfat)`} + \frac{(200 - `r mean(bodyfat$Weight)`)^2}{`r (nrow(bodyfat)-1)*var(bodyfat$Weight)`}}$$
or:
$$(`r predict(bf_wt, newdata = data_frame(Weight=200), interval = "c")[2]`,
`r predict(bf_wt, newdata = data_frame(Weight=200), interval = "c")[3]`)$$


## picture of 95\% CI for mean response - weight model

```{r}
bodyfat %>% 
  ggplot(aes(x=Weight, y=`Pct BF`)) + geom_point(cex=0.5) + geom_smooth(method="lm")
```

## predict $y$ at a new $x$ value

Suppose you want to predict what $y$ might be at some new $x_\nu$ (may
or may not be one of the original $x$'s.)

\pause \textbf{There is no \textit{true} value. We are predicting something random (and un-knowable)---not estimating something fixed (but unknown.)}

\pause What's the "obvious" best guess using the data?
$$\hat y_\nu = b_0 + b_1 x_\nu$$
The *same* guess as the estimate for $\mu_\nu$. 

\pause The variation inherent in such a prediction is different.

## predict a new value---with confidence

A prediction interval will be based on, similar to a confidence interval:
$$\frac{\hat y_\nu - y_\nu}{s.e.(\hat y_\nu - y_\nu)} \sim t_{n-2}$$

\pause The standard error of $\hat y_\nu - y_\nu$ is:
$$s_e\sqrt{1+ \frac{1}{n} + \frac{(x_\nu - \overline x)^2}{S_{xx}}}$$

## predict at `Weight` of 200

\pause The prediction will be (also) `r round(coefficients(bf_wt)[1], 3)` + `r round(coefficients(bf_wt)[2], 3)`(200) = `r predict(bf_wt, newdata = data_frame(Weight=200))`.

\pause But the 95\% "prediction interval" will be quite a bit wider:

$$`r predict(bf_wt, newdata = data_frame(Weight=200))` \pm `r qt(0.975, 248)` \cdot `r glance(bf_wt)$sigma`
\sqrt{1 + \frac{1}{`r nrow(bodyfat)`} + \frac{(200 - `r mean(bodyfat$Weight)`)^2}{`r (nrow(bodyfat)-1)*var(bodyfat$Weight)`}}$$
or:
$$(`r predict(bf_wt, newdata = data_frame(Weight=200), interval = "p")[2]`,
`r predict(bf_wt, newdata = data_frame(Weight=200), interval = "p")[3]`)$$

## picture of 95\% PI weight model

```{r, warning=FALSE}

bf_p <- as_data_frame(predict(bf_wt, interval = "p"))
bf_aug <- cbind(bodyfat, bf_p)

bf_aug %>% 
  ggplot(aes(x=Weight, y=`Pct BF`)) + 
  geom_point() + 
  geom_ribbon(aes(ymin=lwr, ymax=upr), alpha=0.2, linetype=1)
```

## picture of both intervals

```{r}
bf_aug %>% 
  ggplot(aes(x=Weight, y=`Pct BF`)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  geom_ribbon(aes(ymin=lwr, ymax=upr), alpha=0.2, linetype=1)
```

## "you can never be too rich or too thin" - The Duchess of Windsor

```{r}
bodyfat %>% 
  arrange(`Pct BF`)
```

